The {TestIdeals} package for \emph{Macaulay2}

### 2018-10-05 Fsing/TestIdealsPaper/TestIdeals.tex
Formatting consistency changes.

Frobenius roots always exist in polynomial and power series rings, and in $F$-finite regular rings
(cf.~\cite[\S 2]{BlickleMustataSmithDiscretenessAndRationalityOfFThresholds} and \cite[\S 5]{KatzmanParameterTestIdealOfCMRings}).
Below is an example of a computation of a Frobenius root in a polynomial ring.
In Section~\ref{ss: math behind} we describe the main ideas behind such computations.

@@

The generalized Frobenius powers of an ideal are test ideals (see Section~\ref{Section: Test Ideals}) of sufficiently general linear combinations of generators of that ideal.
Thus, the ideals computed above are test ideals of generic quintics in two variables in characteristic $p=3$, and these computations suggest that $\frac35-\frac1{5p^3}$ is a higher $F$-jumping exponent of such quintics, which is indeed the case for any $p\equiv 3 \bmod 5$ (see \cite{hernandez+etal.frobenius_examples}).
To illustrate, compare the computations above with the following.

@@

The calculation above shows that $R/I$ is not $F$-injective (cf. Section \ref{Section: F-singularities}).
Moreover, it shows that $H^2_{\mathfrak{m}} (R/I)\cong \Ann_{E^2} A^t$,
where $E$ is the injective hull of $R_{\mathfrak{m}}/ {\mathfrak{m}} R_{\mathfrak{m}}$,
and the Frobenius map on  $H^2_{\mathfrak{m}} (R/I)$ induced from Frobenius on $R$
is given by $U^t \Theta$, where $\Theta$ is the induced Frobenius on $E$.
The submodule of nilpotent elements in  $H^2_{\mathfrak{m}} (R/I)$ is
given by $\Ann_{E^2} B^t$, where $B$ is the smallest submodule of $R^2$ containing $\Image A + \Image U$ such that $U  B \subseteq B^{[p]}$.
The method \texttt{ascendModule} can be used to calculate  $B$ (see a detailed description of the similar method \texttt{ascendIdeal} in Section \ref{Section: Test Ideals}).

@@

The method \texttt{compatibleIdeals} produces the finite set of $\phi$-compatible prime ideals in the second claim of Theorem \ref{Theorem: finitely many compatible primes}.

@@

\begin{example}
We return to Example \ref{Example: parameter test ideal}.
In (cf.~\cite[\S 9]{KatzmanParameterTestIdealOfCMRings}) it is shown that
there is a surjection $\Ann_E I \rightarrow H^2_{\mathfrak{m}} (R/I)$
which is compatible  with the induced $p^1$-linear map on $H^2_{\mathfrak{m}} (R/I)$
and the $p^1$-linear map $u \Theta$ on $\Ann_E I$, where $u$ is computed as follows.

@@

In \cite{KatzmanZhangAlgorithm}, the algorithms behind the methods in Section \ref{Section: compatible ideals} were extended
to compute prime annihilators of submodules of Artinian modules compatible with a given $p^{e}$-linear map.
This would require, among other things, a faster implementation of our method for finding Frobenius roots of submodules of free modules.

>>>>

Frobenius roots always exist in polynomial and power series rings, and in $F$-finite regular rings
(cf.~\cite[\S 2]{BlickleMustataSmithDiscretenessAndRationalityOfFThresholds} and \cite[\S 5]{KatzmanParameterTestIdealOfCMRings}).
Below is an example of a computation of a Frobenius root in a polynomial ring.
In \autoref{ss: math behind} we describe the main ideas behind such computations.

@@

The generalized Frobenius powers of an ideal are test ideals (see \autoref{Section: Test Ideals}) of sufficiently general linear combinations of generators of that ideal.
Thus, the ideals computed above are test ideals of generic quintics in two variables in characteristic $p=3$, and these computations suggest that $\frac35-\frac1{5p^3}$ is a higher $F$-jumping exponent of such quintics, which is indeed the case for any $p\equiv 3 \bmod 5$ (see \cite{hernandez+etal.frobenius_examples}).
To illustrate, compare the computations above with the following.

@@

The calculation above shows that $R/I$ is not $F$-injective (cf. \autoref{Section: F-singularities}).
Moreover, it shows that $H^2_{\mathfrak{m}} (R/I)\cong \Ann_{E^2} A^t$,
where $E$ is the injective hull of $R_{\mathfrak{m}}/ {\mathfrak{m}} R_{\mathfrak{m}}$,
and the Frobenius map on  $H^2_{\mathfrak{m}} (R/I)$ induced from Frobenius on $R$
is given by $U^t \Theta$, where $\Theta$ is the induced Frobenius on $E$.
The submodule of nilpotent elements in  $H^2_{\mathfrak{m}} (R/I)$ is
given by $\Ann_{E^2} B^t$, where $B$ is the smallest submodule of $R^2$ containing $\Image A + \Image U$ such that $U  B \subseteq B^{[p]}$.
The method \texttt{ascendModule} can be used to calculate  $B$ (see a detailed description of the similar method \texttt{ascendIdeal} in \autoref{Section: Test Ideals}).

@@

The method \texttt{compatibleIdeals} produces the finite set of $\phi$-compatible prime ideals in the second claim of \autoref{Theorem: finitely many compatible primes}.

@@

\begin{example}
We return to \autoref{Example: parameter test ideal}.
In (cf.~\cite[\S 9]{KatzmanParameterTestIdealOfCMRings}) it is shown that
there is a surjection $\Ann_E I \rightarrow H^2_{\mathfrak{m}} (R/I)$
which is compatible  with the induced $p^1$-linear map on $H^2_{\mathfrak{m}} (R/I)$
and the $p^1$-linear map $u \Theta$ on $\Ann_E I$, where $u$ is computed as follows.

@@

In \cite{KatzmanZhangAlgorithm}, the algorithms behind the methods in \autoref{Section: compatible ideals} were extended
to compute prime annihilators of submodules of Artinian modules compatible with a given $p^{e}$-linear map.
This would require, among other things, a faster implementation of our method for finding Frobenius roots of submodules of free modules.
